---
title: "Fitting Models 1 and 2 to data"
author: "Ada Yan"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: TRUE
    keep_tex: yes
---

# Introduction

This document outlines how to fit Models 1 and 2 in "Modelling the viral dynamics of the SARS-CoV-2 Delta and Omicron variants in different cell types" to data.

# Setup

Model fitting is conducted using a Bayesian framework, using Markov Chain Monte Carlo (MCMC) methods implemented in the `lazymcmc` package.

Data and functions used to fit the model to data are in a github repository.  Clone this repository:

```
git clone https://github.com/ada-w-yan/deltaomicronmodelling
```

Also install `lazymcmc`:

```
devtools::install_github("ada-w-yan/lazymcmc")
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

```{r}
# load R package with model fitting functions
git_repo_dir <- "~/git_repos/deltaomicronmodelling/" # change this to the directory of the git repository
devtools::load_all(git_repo_dir)
setwd(git_repo_dir)
library(lazymcmc)
library(parallel) # for parallel processing
```

# Data

The data are in the `data/` folder in the `deltaomicronmodelling` repository.  Note that `V` denotes infectious virus as measured by plaque assay and `V_tot` denotes total virus as measured by RT-qPCR.

# Model 1

The below code fits Model 1 to the data, for hNECs and Calu-3 cells separately.  It outputs to the `model_1_results/hNEC/` and `model_1_results/Calu3/` folders:

- `1_prctile_combined.csv` which contains draws from the posterior distribution;
- `1_summary_chain_combined.csv` which, in addition to draws from the posterior distribution, contains summary statistics ($R_0$, growth rate and doubling time)
- `model_predictions_omicron.png` etc. Figures 5 and 6: Model 1 fits to data. Dots show the data, black lines show the maximum likelihood
fit, shaded areas show the 95% credible interval (CrI) and dotted lines show the limit of detection.

```{r}
dir.create("model_1_results/")

run_model_1_and_postprocess <- function(cells) {
  dir_name <- paste0("model_1_results/", cells, "/")
  # fit model to data
  run_exp_two_strain(save_dir = dir_name,
                     cells == "Calu3",
                     length_run = 2,
                     run_flag = TRUE)
  
  trajectory_filename <- paste0(dir_name, "trajectories.rds")
  chain_filename <- paste0(dir_name, "1.RData")
  # calculate median and 95% CI for model trajectories
  trajectories <- calc_trajectories(chain_filename,
                                    "model_teiv_dual.R",
                                    two_strain = TRUE)
  saveRDS(trajectories, trajectory_filename)
  
  # plot median and 95% CI for model trajectories
  lapply(c("omicron", "delta"),
         \(x) plot_trajectories(x, cells == "Calu3", trajectory_filename))
}

lapply(c("hNEC", "Calu3"), run_model_1_and_postprocess)
```

# Model 2

The below code fits Model 1 to the data, for hNECs and Calu-3 cells separately.  It outputs to e.g. `model_2_results/omicron_hNEC/` folders:

- `1_prctile_combined.csv` which contains draws from the posterior distribution;
- `1_summary_chain_combined.csv` which, in addition to draws from the posterior distribution, contains summary statistics ($R_0$, growth rate and doubling time, for all combinations of pathways and with and without endosomal restriction)
- `model_predictions_omicron.png` etc. Figures 78 and : Model 2 fits to data. Dots show the data, black lines show the maximum likelihood
fit, shaded areas show the 95% credible interval (CrI) and dotted lines show the limit of detection.

```{r}
run_model_2_and_postprocess <- function(strain, cells) {
  
  dir_name <- paste0("model_2_results/", strain, "_", cells, "/")
  
  # fit model to data
  run_exp_camostat_amphoB(save_dir = dir_name,
                          strain = strain,
                          Calu3 = (cells == "Calu3"),
                          length_run = 2,
                          run_flag = TRUE)
  
  trajectory_filename <- paste0(dir_name, "trajectories.rds")
  chain_filename <- paste0(dir_name, "1.RData")
  
  # calculate median and 95% CI for model trajectories
  trajectories <- calc_trajectories(chain_filename,
                                    "model_different_eclipse_dual.R",
                                    two_strain = FALSE)
  saveRDS(trajectories, trajectory_filename)
  
  # plot median and 95% CI for model trajectories
  plot_trajectories_camostat_amphoB(strain, cells == "Calu3", trajectory_filename)
}

# run for each combination of virus strain and cell type
par_grid <- expand_grid(strain = c("omicron", "delta"), cells = c("hNEC", "Calu3"))
apply_named_args(par_grid, 1, run_model_2_and_postprocess)
```

# Sensitivity analysis

Changing $\tau_E$ and $\tau_T$ for model 2:

Note this does not make the plots, only calculates summary statistics.

```{r}
run_sensitivity <- function(strain, cells, tau_E, tau_T) {
  dir_name <- paste0("sens/", strain, "_", cells, "_k_E_", tau_E, "_k_T_", tau_T, "/")
  # fit model to data
  run_exp_camostat_amphoB(save_dir = dir_name,
                          strain = strain,
                          Calu3 = (cells == "Calu3"),
                          tau_E = tau_E,
                          tau_T = tau_T,
                          length_run = 2,
                          run_flag = TRUE)
}

par_grid <- expand_grid(strain = c("omicron", "delta"),
                        cells = c("hNEC", "Calu3"),
                        tau_E = c(1,2,4),
                        tau_T = c(2,4,6))

apply_named_args(par_grid, 1, run_sensitivity)
```

```{r}
```

